### 2024-10-09 esop25/cpj.tex
checkpoint

For example, given some string $y$ and a message send such as:
$$
\xassign{\mx{w}{1}}{\secret{w} \ftimes \flip{w}}{2}
$$
we would assign the following type to denote that $\mx{w}{1}$ is dependent on
both $\sx{w}{2}$ and $\rx{w}{2}$:
$$
\mx{w}{1} : \setit{\sx{w}{2}, \rx{w}{2}}
$$
But in the case of 2-party reconstructive sharing:
$$
\xassign{\mx{w}{1}}{\secret{w} \fminus \flip{w}}{2}
$$
we want the type of $\mx{w}{1}$ to be an ``encrypted'' type
that is by itself independent of $\sx{w}{2}$. However, we also
want to track the dependence of $\mx{w}{2}$ and $\rx{w}{2}$.
In short, we assign $\mx{w}{1}$ a ``ciphertext type'':
$$
\mx{w}{1} : \cty{\rx{w}{2}}{\setit{\sx{w}{2}}}
$$
indicating syntactically that $\mx{w}{1}$ is an encryption of a
value dependent on $\sx{w}{2}$ with $\rx{w}{2}$ and as long as client
1 does not subsequently receive messages that are dependent on
$\rx{w}{2}$, the value $\sx{w}{2}$ remains hidden.

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and the
``evaluator'' obtains the wire value for its input via OT. The
evaluator shares the initial encrypted wire value of one of its input
secrets as a selection bit-- assuming that client 2 is the garbler and
client 1 is the evaluator, we can write this as:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
This allows us to assign the same ciphertext type to $\mx{w}{1}$ as
above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
equivalence this way has been previously well-studied
\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
it is also important that we ensure that samples are used at most once
for encryption. As in previous work \cite{darais2019language} we can
use a form of type linearity to this end.

@@

The syntax of types $\ty$ and type environments $\Gamma$ are presented
in Figure \ref{fig-cpj}\cnote{need to fix set notations}. We also
define type judgements for expressions and programs by direct
interpretation as terms $\phi$ and constraints $\eqs$ respectively,
via the encoding $\toeq{cdot}$. The \TirName{Encode} rule captures
the use ``fresh'' samples for encryption discussed above. Derivations
enforce linearity by requiring that variables used for encryption
are added to judgements, and guaranteed to not be used elsewhere
via a form of union that ensures disjointness of combined sets. 
\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ iff $R_1 \cap R_2 = \varnothing$.
\end{definition}
The \TirName{DepTy} rule handles all other syntactic forms and just
records the dependencies on variables in the term.  The \TirName{Send}
rule records the types of messages, reveals, and outputs in
enviroments, and the \TirName{Seq} rule combines the types of
sequenced programs.  Given this we define validity of program type
judgements as follows. \cnote{Define $\eqspre$.}
\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}
It is important to note that this presentation of derivations is
logical, not algorithmic-- in particular it is not syntax-directed
due to the \TirName{Encode} rule.

>>>>

For example, given some string $y$ and a message send such as:
$$
\xassign{\mx{w}{1}}{\secret{w} \ftimes \flip{w}}{2}
$$
we would assign the following type to denote that $\mx{w}{1}$ is dependent on
both $\sx{w}{2}$ and $\rx{w}{2}$:
$$
\mx{w}{1} : \setit{\sx{w}{2}, \rx{w}{2}}
$$
But in the case of 2-party reconstructive sharing:
$$
\xassign{\mx{w}{1}}{\secret{w} \fminus \flip{w}}{2}
$$
we want the type of $\mx{w}{1}$ to be an ``encrypted'' type
that is by itself independent of $\sx{w}{2}$. However, we also
want to track the dependence of $\mx{w}{2}$ and $\rx{w}{2}$.
In short, we assign $\mx{w}{1}$ a ``ciphertext type'', aka
a \emph{ciphertype}:
$$
\mx{w}{1} : \cty{\rx{w}{2}}{\setit{\sx{w}{2}}}
$$
indicating syntactically that $\mx{w}{1}$ is an encryption of a
value dependent on $\sx{w}{2}$ with $\rx{w}{2}$ and as long as client
1 does not subsequently receive messages that are dependent on
$\rx{w}{2}$, the value $\sx{w}{2}$ remains hidden.

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and the
``evaluator'' obtains the wire value for its input via OT. The
evaluator shares the initial encrypted wire value of one of its input
secrets as a selection bit-- assuming that client 2 is the garbler and
client 1 is the evaluator, we can write this as:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
This allows us to assign the same ciphertype to $\mx{w}{1}$ as
above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
equivalence this way has been previously well-studied
\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
it is also important that we ensure that samples are used at most once
for encryption. As in previous work \cite{darais2019language} we can
use a form of type linearity to this end.

@@

The syntax of types $\ty$ and type environments $\Gamma$ are presented
in Figure \ref{fig-cpj}\cnote{need to fix set notations}. We also
define type judgements for expressions and programs by direct
interpretation as terms $\phi$ and constraints $\eqs$ respectively,
via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule captures
the use ``fresh'' samples for encryption discussed above. Derivations
enforce linearity by requiring that variables used for encryption
are added to judgements, and guaranteed to not be used elsewhere
via a form of union that ensures disjointness of combined sets. 
\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ iff $R_1 \cap R_2 = \varnothing$.
\end{definition}
The \TirName{DepTy} rule handles all other syntactic forms and just
records the dependencies on variables in the term.  The \TirName{Send}
rule records the types of messages, reveals, and outputs in
enviroments, and the \TirName{Seq} rule combines the types of
sequenced programs.  Given this we define validity of program type
judgements as follows. \cnote{Define $\eqspre$.}
\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}
It is important to note that this presentation of derivations is
logical, not algorithmic-- in particular it is not syntax-directed
due to the \TirName{Encode} rule.

### 2024-10-09 esop25/examples.tex
checkpoint

@@

\begin{verbatimtab}
andtableygc(g,x,y)
{
   let table = (~r[g],~r[g],~r[g],r[g])
   in permute4(r[x],r[y],table)
}

>>>>

We consider several examples that illustrate the versatility of the $\TirName{Encode}$
typing rule for confidentiality, the use of postconditions for correctness,
and the use of both pre-~and postconditions for integrity.

@@

The $\ttt{encode}$ function example in Section \ref{section-metalangty}
illustrates the use of type annotations to track confidentiality. In
2-party Yao's Garbled Circuits (YGC), a \emph{garbler} prepares a circuit encoding
that is shared with a \emph{evaluator}, and the $\ttt{encode}$ function
defines the method the garbler uses to share their input secret encodings
with the evaluator. The example illustrates how confidentiality is reflected
in its (cipher)type.

In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
binary circuit protocol, values $v$ are encoded as reconstructive
shares $v_1$ and $v_2$ held by each party with $v = v_1 \fplus v_2$.
We will follow the convention that shared values are identified by
strings $w$ and encoded as shares $\mx{w}{1}$ and $\mx{w}{2}$.  As for
YGC, ciphertypes reflect the confidentiality of GMW input encodings as
defined in the $\ttt{encodegmw}$ function in Figure \ref{fig-gmw}. No
programmer annotation is needed given the syntax of the function
body. More interestingly, the $\ttt{andgmw}$ gate definition shows how
a programmer hint can express the relevant confidentiality property of
the output share $\mx{z}{2}$ on client 2. The non-trivial equivalence
can be verified by SMT once during verification, and subsequently
confidentiality is expressed in the dependent type of the function
(as for $\ttt{encode}$).

Note also that the $\ttt{andgmw}$ function is decorated with a
postcondition that expresses the correctness property of the
function. Although not strictly necessary for confidentiality, this
annotation can at least help eliminate bugs, and also can be used
compositionally for whole-program correctness properties. As for
confidentiality hints, this postcondition needs to be verified only
once at the point of definition and subsequently is guaranteed to hold
for any application.

\input{gmw}

In the 2-party GMW protocol \cite{evans2018pragmatic}, another garbled
binary circuit protocol, values $v$ are encoded as reconstructive
shares $v_1$ and $v_2$ held by each party with $v = v_1 \fplus v_2$.
We will follow the convention that shared values are identified by
strings $w$ and encoded as shares $\mx{w}{1}$ and $\mx{w}{2}$.  As for
YGC, ciphertypes reflect the confidentiality of GMW input encodings as
defined in the $\ttt{encodegmw}$ function in Figure \ref{fig-gmw}. No
programmer annotation is needed given the syntax of the function
body. More interestingly, the $\ttt{andgmw}$ gate definition shows how
a programmer hint can express the relevant confidentiality property of
the output share $\mx{z}{2}$ on client 2. The non-trivial equivalence
can be verified by SMT once during verification, and subsequently
confidentiality is expressed in the dependent type of the function
(as for $\ttt{encode}$).

Note also that the $\ttt{andgmw}$ function is decorated with a
postcondition that expresses the correctness property of the
function. Although not strictly necessary for confidentiality, this
annotation can at least help eliminate bugs, and also can be used
compositionally for whole-program correctness properties. As for
confidentiality hints, this postcondition needs to be verified only
once at the point of definition and subsequently is guaranteed to hold
for any application.

\subsubsection{Integrity Examples}

\input{bdoz}

We can capture both the preliminary distribution of Beaver triples and
BDOZ shares as a pre-processing predicate that establishes conditions
for initial memories (see Definition \ref{def-aprogtt}).  Here we
assume two input secrets $\elab{\secret{x}}{1}$ and
$\elab{\secret{y}}{2}$ and a single Beaver Triple to compute
$\elab{\secret{x}}{1} \ftimes \elab{\secret{y}}{2}$, but we can extend
this for additional gates.  As for GMW, we use
$\macgv{\mesg{w}}$\cnote{need to update this} to refer to
secret-shared values reconstructed with addition, but for BDOZ by
convention each share of $\macgv{\mesg{w}}$ is represented as
$\elab{\mesg{w\ttt{s}}}{\cid}$, the MAC of which is represented as a
$\elab{\mesg{w\ttt{m}}}{\cid}$ for all $\cid$, and each client holds a
key $\elab{\mesg{w\ttt{k}}}{\cid}$ for authentication of the other's
share.
$$
\begin{array}{l}
\mx{\ttt{"xs"}}{2} \eop \sx{\ttt{"x"}}{1} \fplus \rx{\ttt{"x"}}{1} \wedge 
\mx{\ttt{"xs"}}{1} \eop \rx{\ttt{"x"}}{1} \wedge \\
\mx{\ttt{"ys"}}{1} \eop \sx{\ttt{"y"}}{2} \fplus \rx{\ttt{"y"}}{2} \wedge 
\mx{\ttt{"ys"}}{2} \eop \rx{\ttt{"y"}}{2} 
\end{array}
$$

$$
\begin{array}{l}
\mx{\ttt{"delta"}}{1} \eop \rx{\ttt{"delta"}}{1} \wedge
\mx{\ttt{"xk"}}{1} \eop \rx{\ttt{"xk"}}{1} \wedge\\
\mx{\ttt{"xm"}}{2} \eop \mx{\ttt{"xk"}}{1} \fplus (\mx{\ttt{"delta"}}{1} * \mx{"xs"}{2})
\end{array}
$$

$$
\begin{array}{l}
\mx{\ttt{"g1as"}}{1} \eop \rx{\ttt{"g1as"}}{1} \wedge 
\mx{\ttt{"g1bs"}}{1} \eop \rx{\ttt{"g1bs"}}{1} \wedge\\
\mx{\ttt{"g1as"}}{2} \eop \rx{\ttt{"g1as"}}{2} \wedge 
\mx{\ttt{"g1bs"}}{2} \eop \rx{\ttt{"g1bs"}}{2} \wedge\\
\mx{\ttt{"g1cs"}}{1} \fplus \mx{\ttt{"g1cs"}}{2} \eop \\
\quad (\mx{\ttt{"g1as"}}{1} \fplus \mx{\ttt{"g1bs"}}{2}) \ftimes
      (\mx{\ttt{"g1bs"}}{1} \fplus \mx{\ttt{"g1bs"}}{2}) 
\end{array}
$$

%%%%%%%%%%%%%%%%%%
%%% ARCHIVAL BELOW
%%%%%%%%%%%%%%%%%%
\begin{comment}
\begin{verbatimtab}
andtableygc(g,x,y)
{
   let table = (~r[g],~r[g],~r[g],r[g])
   in permute4(r[x],r[y],table)
}

### 2024-10-09 esop25/ipj.tex
checkpoint

As in \cite{skalka-near-ppdp24}, we use $\macgv{\mesg{w}}$ to refer to
secret-shared values reconstructed with addition, and by
convention each share of $\macgv{\mesg{w}}$ is represented as
$\elab{\mesg{w\ttt{s}}}{\cid}$, the MAC of which is represented as a
$\elab{\mesg{w\ttt{m}}}{\cid}$ for all $\cid$, and each client holds a
key $\elab{\mesg{w\ttt{k}}}{\cid}$ for authentication of the other's
share. Each client also holds a global key $\mesg{\ttt{delta}}$. For
any such share identified by string $w$, the BDOZ MAC scheme is defined 
by the equality predicate $\macbdoz{w}$:
$$
\macbdoz{w} \defeq
\mesg{w\ttt{m}} = \mesg{w\ttt{k}} \fplus \ttt{(}\mesg{\ttt{delta}} \ftimes
\mesg{w\ttt{s}}\ttt{)}
$$

>>>>

As in \cite{skalka-near-ppdp24}, we use $\macgv{\mesg{w}}$ to refer to
secret-shared values reconstructed with addition\cnote{needed?}, and by
convention each share of $\macgv{\mesg{w}}$ is represented as
$\elab{\mesg{w\ttt{s}}}{\cid}$, the MAC of which is represented as a
$\elab{\mesg{w\ttt{m}}}{\cid}$ for all $\cid$, and each client holds a
key $\elab{\mesg{w\ttt{k}}}{\cid}$ for authentication of the other's
share. Each client also holds a global key $\mesg{\ttt{delta}}$. For
any such share identified by string $w$, the BDOZ MAC scheme is defined 
by the equality predicate $\macbdoz{w}$:
$$
\macbdoz{w} \defeq
\mesg{w\ttt{m}} = \mesg{w\ttt{k}} \fplus \ttt{(}\mesg{\ttt{delta}} \ftimes
\mesg{w\ttt{s}}\ttt{)}
$$

### 2024-10-09 esop25/smt.tex
checkpoint

In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
$\eqs$, where constraint terms $\phi$ are similar to expressions $\be$
except the can ``cross party lines''. This turns out to be very useful
to express correctness properties-- for example, in the GMW protocol
wire values in circuits are represented by reconstructive shares
\cite{evans2018pragmatic}.  If by convention shares of values $n$ are
represented by $\mesg{n}$ on clients, then assuming two clients
$\setit{1,2}$ the reconstructed value can be expressed as $\mx{n}{1} +
\mx{n}{2}$.  So, while summing values across clients is disallowed in
$\minicat$ protocols, we can express properties of shared values
via constraints.

@@

\subsubsection{Example: Correctness of 3-Party Addition}
Consider the following 3-party Shamir Addition protocol.

>>>>

In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
$\eqs$, where constraint terms $\phi$ are similar to expressions $\be$
except the can ``cross party lines''. This turns out to be very useful
to express correctness properties-- for example, in the
Goldreich-Micali-Wigderson (GMW) protocol wire values in circuits are
represented by reconstructive shares \cite{evans2018pragmatic}.  If by
convention shares of values $n$ are represented by $\mesg{n}$ on
clients, then assuming two clients $\setit{1,2}$ the reconstructed
value can be expressed as $\mx{n}{1} + \mx{n}{2}$.  So, while summing
values across clients is disallowed in $\minicat$ protocols, we can
express properties of shared values via constraints.

@@

\subsubsection{Example: Correctness of 3-Party Addition}
Consider the following 3-party additive secret sharing protocol.
